Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-27 paper/paper.tex
merged final in old style

\documentclass[sigconf]{acmart}

\setcopyright{rightsretained}

\usepackage{siunitx}
\usepackage[color]{changebar}
\usepackage{enumitem}
\usepackage[lighttt]{lmodern}
\usepackage{graphicx}

\usepackage{listings}

\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}
\lstdefinestyle{default}{language=Anglican,basicstyle=\ttfamily, columns=flexible, showstringspaces=false}

\usepackage{url}
\usepackage{etoolbox}

@@

@@

\acmConference[IFL'16]{Implementation and Application of Functional Programming Languages}{August 2016}{Leuven, Belgium}
\acmYear{2016}
\copyrightyear{2017}

\begin{document}
\title{Design and Implementation of Probabilistic Programming Language Anglican}

\author{David Tolpin}
\affiliation{\institution{University of Oxford}}
\email{dtolpin@robots.ox.ac.uk}

\author{Jan-Willem van de Meent}
\affiliation{\institution{University of Oxford}}
\email{jwvdm@robots.ox.ac.uk}

\author{Hongseok Yang}
\affiliation{\institution{University of Oxford}}
\email{hongseok.yang@cs.ox.ac.uk}

\author{Frank Wood}
\affiliation{\institution{University of Oxford}}
\email{fwood@cs.ox.ac.uk}

@@

Probabilistic programming systems \cite{GMR+08,MSP14,WVM14,GS15}
represent generative models as programs written in a specialized
language that provides syntax for the definition and conditioning of
random variables. The code for such models is generally concise,
modular, and easy to modify or extend. Typically inference can be
performed for any probabilistic program using one or more generic
inference techniques provided by the system back end, such as
Metropolis-Hastings \cite{WSG11,MSP14,YHG14}, Hamiltonian Monte Carlo
\cite{SDT14}, expectation propagation \cite{MWG+10}, and extensions
of Sequential Monte Carlo \cite{WVM14,MYM+15,PWD+14} methods.
These generic techniques may be less statistically efficient
than techniques tailored to a specific model. However,
probabilistic programming facilitates simpler implementations of
models making the inference inherently faster.
\jw{I'm not sure I understand how PP facilitates simpler implementations that make inference inherently faster.}

@@

@@

\begin{figure}
    \centering
    \includegraphics[trim={12pt 0 0 0},scale=0.6]{same-customer-time.pdf}
    \caption{Arrival time distribution for a single customer.}
    \label{fig:dist-same}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[trim={12pt 0 0 0},scale=0.6]{different-customers-times.pdf}
    \caption{Arrival time distributions for two customers.}
    \label{fig:dist-diff}
\end{figure}

@@

;; invoked when algo parameter is :importance
(defmethod infer :importance 
  [algo prog value & {}]
  (letfn
    ;; recursive function without any parameter
    [sample-seq 
      []
      ;; lazy infinite sequence 
      (lazy-seq
        (cons
          (:state (exec ::algorithm prog value
                        initial-state))
          (sample-seq))))]
    (sample-seq)))
\end{lstlisting}
This implementation is dispatched 
when \texttt{infer} receives {\tt :im\-por\-tan\-ce} as its \texttt{algo} 
parameter. (Anglican includes multiple implementations of \texttt{infer}
for different inference algorithms.) Once dispatched, it lazily 
constructs an infinite sequence of inference states by repeatedly 
executing the program \texttt{prog} using \texttt{exec}
and retrieving the final inference state of the execution 
using \texttt{:state}. For checkpoint handlers, importance sampling simply
relies on their default implementations.

@@

\section*{Acknowledgements}

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

>>>>

\documentclass[preprint]{sigplanconf}

\usepackage{siunitx}
\usepackage[color]{changebar}
\usepackage{enumitem}
\usepackage[lighttt]{lmodern}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
%\usepackage{microtype}

\usepackage{listings}
\usepackage{lmodern}
\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}
\lstdefinestyle{default}{language=Anglican,basicstyle=\linespread{1.0}\ttfamily, columns=flexible, showstringspaces=false}

\usepackage{hyperref}
\usepackage{url}
\usepackage{etoolbox}

@@

\title{Design and Implementation of Probabilistic Programming Language Anglican}
\authorinfo{David Tolpin \and Jan-Willem van de Meent \and Hongseok Yang \and Frank Wood}{University of Oxford}{dtolpin@robots.ox.ac.uk \and jwvdm@robots.ox.ac.uk \and hongseok.yang@cs.ox.ac.uk \and fwood@robots.ox.ac.uk}

@@

\begin{document}

@@

Probabilistic programming systems \cite{GMR+08,MSP14,WVM14,GS15}
represent generative models as programs written in a specialized
language that provides syntax for the definition and conditioning of
random variables. The code for such models is generally concise,
modular, and easy to modify or extend. Typically inference can be
performed for any probabilistic program using one or more generic
inference techniques provided by the system back end, such as
Metropolis-Hastings \cite{WSG11,MSP14,YHG14}, Hamiltonian Monte Carlo
\cite{SDT14}, expectation propagation \cite{MWG+10}, and extensions
of Sequential Monte Carlo \cite{WVM14,MYM+15,PWD+14} methods.

@@

%Their operational semantics can be defined in terms of
%importance sampling(Equation~\ref{eqn:p-trace}): 
%\begin{itemize}
%    \item \texttt{sample} returns a value drawn from
%        the distribution.
%    \item \texttt{observe} multiplies the unnormalized
%        probability of the program trace by the probability of
%        drawing the value from the distribution.
%\end{itemize}
%However, the actual computation involved in handling of
%\texttt{sample} and \texttt{observe} depends on the inference
%algorithm. Different inference algorithms may treat
%\texttt{sample} and \texttt{observe} differently as long as they
%compute the same distribution of program results.
%Denotational and operational semantics for an idealized version
%of Anglican is introduced in \cite{SYH+16}; in general though,
%specifying semantics of higher-order probabilistic programming
%languages is an open problem.

@@

\begin{figure}
    \centering
    \includegraphics[trim={12pt 0 0 0},scale=0.65]{same-customer-time.pdf}
    \caption{Arrival time distribution for a single customer.}
    \label{fig:dist-same}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[trim={12pt 0 0 0},scale=0.65]{different-customers-times.pdf}
    \caption{Arrival time distributions for two customers.}
    \label{fig:dist-diff}
\end{figure}

@@

;; invoked when algo parameter is :importance
(defmethod infer :importance 
  [algo prog value & {}]
  (letfn
    ;; recursive function without any parameter
    [sample-seq 
      []
      ;; lazy infinite sequence 
      (lazy-seq
        (cons
          (:state (exec ::algorithm prog value
                        initial-state))
          (sample-seq))))]
    (sample-seq)))
\end{lstlisting}
This implementation is dispatched 
when \texttt{infer} receives \texttt{:importance} as its \texttt{algo} 
parameter. (Anglican includes multiple implementations of \texttt{infer}
for different inference algorithms.) Once dispatched, it lazily 
constructs an infinite sequence of inference states by repeatedly 
executing the program \texttt{prog} using \texttt{exec}
and retrieving the final inference state of the execution 
using \texttt{:state}. For checkpoint handlers, importance sampling simply
relies on their default implementations.

@@

\acks

@@

\bibliographystyle{abbrvnat}
\bibliography{refs}
